Supplementary Materials- A Bayesian approach to burial networks to infer social changes in northeastern Taiwan during the European colonization period

### 2021-01-26 analysis/paper/supplementary-material.Rmd
edit SI
showing visual md editing

## Kiwulan in the 17th century northern Taiwan {-}

<!--fit in a bigger historical picture in a periclonial context: move to SI -->
The involvement in long-distance exchange is often associated with changes in political-economic strategies to a network mode, where ambitious individuals are able to build their local power base and personal influence through the distribution of high-value goods [@Blanton1996; @Feinman2000; @Klehm2017]. Northeastern Taiwan was involved in a regional exchange in East Asia since the 14th century and long-distance trade during the European presence in the 17th century [@Chen2005; @Wang2007]. In 1626, the Spanish built their forts in Heping dao and Tamsui, northern Taiwan, and later were defeated and taken over by the Dutch in 1642 [@Andrade2007]. Indigenous people in northeastern Taiwan first encountered colonial power in 1632 when they were attacked by the Spanish and later in 1647 by the Dutch, who also asked them to pay annual tributes in a flexible way [@Borao2009; @Kang2016]. The Dutch would give local Indigenous leaders symbolic items or trade goods to assert their political authority and consolidate their relationship with Indigenous societies.

@@

<!-- burial data and the definition of tie formation: partially move to SI  --> 
For trade beads with substantial differences in quantities between burials, we described each burial as having one of four levels according to their distributions across all burials (Figure \@ref(fig:figure-raincloud-beads)). Gold-foil beads are in levels of high (>3), upper-middle (3), lower-middle (2), and low (1); carnelian beads are in levels of high (>6), upper-middle (4-6), lower-middle (3), and low (1-2); glass beads are in levels of high (>6), upper-middle (3-6), lower-middle (2), and low (1). If burial 1 and burial 2 both have high quantities of agate beads, then there will be a tie connecting them.

@@

## Review of network analysis and archaeological applications {-}

<!-- intro to networks: move to SI -->
A social network is generally visualized as a graph consisting of a set of socially-relevant nodes/actors, connected by edges/ties representing one or more relations, such as friendship, collaborations, information flow, trade ties, or any other forms of connection of interest [@Marin2011; @Wasserman1994]. The ties can be classified into four major types, including similarities, social relations, interactions, and flows [@Borgatti2009]. In archaeology, actors can be people, groups, objects, places, or events, with ties built on similarity, proximity, or co-presence of material culture to create patterns reflecting influence, geographical distance, or affiliations in social groups [@Mills2017; @Brughmans2018; @Peeples2019]. For example, past trade can be conceptualized as a network of individual entities connected by shared similarities, the flow of goods, to represent their interactions [@Collar2015, pp. 4]. Similar concepts can be applied to a wide range of archaeological data with relational assumptions, such as burial contexts. Burial goods, especially high value goods, can reflect social practices in broader cultural contexts to represent personal wealth or social status from which we can infer social differentiation or complexity [@Gamble2002; @Janes2013]. This enables the exploration of the structure of the past social organization through the identification of the relationships among burials.

<!-- two types of network analysis: move to SI -->
Network analysis has been increasingly applied by archaeologists in recent years to deal with past interactions and explore the underlying mechanisms. There are two common approaches to characterize network properties at two distinct scales: node/edge level and graph level [@Peeples2019]. Node level focuses on the role of node in a network, such as centrality, representing the individual influence or social prominence in a group, while graph level assesses the whole network attributes, such as density, clustering in a network, to generalize relationship patterns [@Mills2017; @Peeples2019]. By quantifying those network properties, archaeologists can answer a wide range of research questions.

@@

<!-- current approaches to network analysis: move to SI -->
Current approaches to network analysis used by archaeologists are mostly restricted to a single rational structure without consideration of interaction between network variables. Our use of Bayesian inference on ERGMs is the first application to archaeological data that can bring new insights to understand past social structures by characterizing network properties as a whole. ERGMs are an important family of statistical models for networks that allows direct modeling for the formation of edges, or ties, between nodes [@Robins2007].

@@

## Further details on methods of modeling fitting {-}

@@

<!--Model fitting: move to SI --> 
After we set our model parameters, we simulated networks in a Bayesian framework using a Markov chain Monte Carlo (MCMC) algorithm. MCMC algorithms allow estimation of posterior distributions through direct random sampling the posterior without assuming the prior comes from any specific distribution [@Hamra2013]. We can obtain a posterior distribution by constructing a Markov chain that describes a sequence of moves from current state to the next state following probabilistic rules based on an algorithm. This enables a random or stochastic simulation in a long run where each move does not depend on the previous move. More chains ensures a more desirable posterior distribution that is close to the target distribution under study, or convergence.

@@

# References {-}

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->
<div id="refs"></div>

>>>>

## Kiwulan in the 17th century northern Taiwan {.unnumbered}

<!--fit in a bigger historical picture in a periclonial context: move to SI -->

The involvement in long-distance exchange is often associated with changes in political-economic strategies to a network mode, where ambitious individuals are able to build their local power base and personal influence through the distribution of high-value goods [@Blanton1996; @Feinman2000; @Klehm2017]. Northeastern Taiwan was involved in a regional exchange in East Asia since the 14th century and long-distance trade during the European presence in the 17th century [@Chen2005; @Wang2007]. In 1626, the Spanish built their forts in Heping dao and Tamsui, northern Taiwan, and later were defeated and taken over by the Dutch in 1642 [@Andrade2007]. Indigenous people in northeastern Taiwan first encountered colonial power in 1632 when they were attacked by the Spanish and later in 1647 by the Dutch, who also asked them to pay annual tributes in a flexible way [@Borao2009; @Kang2016]. The Dutch would give local Indigenous leaders symbolic items or trade goods to assert their political authority and consolidate their relationship with Indigenous societies.

@@

<!-- burial data and the definition of tie formation: partially move to SI  -->

For trade beads with substantial differences in quantities between burials, we described each burial as having one of four levels according to their distributions across all burials (Figure \@ref(fig:figure-raincloud-beads)). Gold-foil beads are in levels of high (\>3), upper-middle (3), lower-middle (2), and low (1); carnelian beads are in levels of high (\>6), upper-middle (4-6), lower-middle (3), and low (1-2); glass beads are in levels of high (\>6), upper-middle (3-6), lower-middle (2), and low (1). If burial 1 and burial 2 both have high quantities of agate beads, then there will be a tie connecting them.

@@

## Review of network analysis and archaeological applications {.unnumbered}

<!-- intro to networks: move to SI -->

A social network is generally visualized as a graph consisting of a set of socially-relevant nodes/actors, connected by edges/ties representing one or more relations, such as friendship, collaborations, information flow, trade ties, or any other forms of connection of interest [@Marin2011; @Wasserman1994]. The ties can be classified into four major types, including similarities, social relations, interactions, and flows [@Borgatti2009]. In archaeology, actors can be people, groups, objects, places, or events, with ties built on similarity, proximity, or co-presence of material culture to create patterns reflecting influence, geographical distance, or affiliations in social groups [@Mills2017; @Brughmans2018; @Peeples2019]. For example, past trade can be conceptualized as a network of individual entities connected by shared similarities, the flow of goods, to represent their interactions [@Collar2015, pp. 4]. Similar concepts can be applied to a wide range of archaeological data with relational assumptions, such as burial contexts. Burial goods, especially high value goods, can reflect social practices in broader cultural contexts to represent personal wealth or social status from which we can infer social differentiation or complexity [@Gamble2002; @Janes2013]. This enables the exploration of the structure of the past social organization through the identification of the relationships among burials.

<!-- two types of network analysis: move to SI -->

Network analysis has been increasingly applied by archaeologists in recent years to deal with past interactions and explore the underlying mechanisms. There are two common approaches to characterize network properties at two distinct scales: node/edge level and graph level [@Peeples2019]. Node level focuses on the role of node in a network, such as centrality, representing the individual influence or social prominence in a group, while graph level assesses the whole network attributes, such as density, clustering in a network, to generalize relationship patterns [@Mills2017; @Peeples2019]. By quantifying those network properties, archaeologists can answer a wide range of research questions.

@@

<!-- current approaches to network analysis: move to SI -->

Current approaches to network analysis used by archaeologists are mostly restricted to a single rational structure without consideration of interaction between network variables. Our use of Bayesian inference on ERGMs is the first application to archaeological data that can bring new insights to understand past social structures by characterizing network properties as a whole. ERGMs are an important family of statistical models for networks that allows direct modeling for the formation of edges, or ties, between nodes [@Robins2007].

@@

## Further details on methods of modeling fitting {.unnumbered}

@@

<!--Model fitting: move to SI -->

After we set our model parameters, we simulated networks in a Bayesian framework using a Markov chain Monte Carlo (MCMC) algorithm. MCMC algorithms allow estimation of posterior distributions through direct random sampling the posterior without assuming the prior comes from any specific distribution [@Hamra2013]. We can obtain a posterior distribution by constructing a Markov chain that describes a sequence of moves from current state to the next state following probabilistic rules based on an algorithm. This enables a random or stochastic simulation in a long run where each move does not depend on the previous move. More chains ensures a more desirable posterior distribution that is close to the target distribution under study, or convergence.

@@

# References {.unnumbered}

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

::: {#refs}
:::
